---
title: "Exploring brick experiment data"
author: "Ingrid Knapp"
date: "10th August, 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse) #install.packages(tidyverse)
library(janitor)
library(ggplot2)
library(RColorBrewer)
library(dbplyr)
#check your working directory
getwd()
#setwd if need to
# https://github.com/ingridknapp/himb_coral_nursery
#setwd()

```
  
  
```{r import_brick_experiment_data, echo=TRUE, message=FALSE}

data <- read.csv("https://raw.githubusercontent.com/ingridknapp/himb_coral_nursery/master/R_Brick_array_data_2017-18_Montipora_v1.4.csv")

```
  
  
```{r tidying, echo=TRUE, message=FALSE}


#have a look at the data
glimpse(data)


#change integers into factors for columns that need changing

data$genotype = as.factor(data$genotype)
data$brick_colour = as.factor(data$brick_colour)

#clean up the headings
data<- janitor::clean_names(data)

#have a look at the data
glimpse(data)
head(data)




```
```{r means_&_SE, echo=TRUE, message=FALSE}
#USe dplyr to aggregate and calculate means, standard deviations, n and standard errors

#array (A,B,C,D)
array_summarise<-data %>%
  group_by(array) %>%
  summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE), n=n(),se=sd/sqrt(n))

#genotype (1,2,3)
genotype_summarise<-data %>%
  group_by(genotype) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))

#brick side (Top, N, S, E, W)
side_summarise<-data %>%
  group_by(side) %>%
   summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))

#brick_colour (1=white, 2=grey, 3=pink)
brick_colour_summarise<-data %>%
  group_by(brick_colour) %>%
  summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))

#brick_id (various)
brick_id_summarise<-data %>%
  group_by(brick_id) %>%
  summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))


# view all summary tables

array_summarise
side_summarise
genotype_summarise
brick_colour_summarise
brick_id_summarise




```
  
  
  
```{r graphs, echo=TRUE, message=FALSE}

#Basic plots looking at means and SE for each factor

#array plot (A,B,C,D)
array_plot <- ggplot (data=array_summarise, aes(x=array, y=mean, fill=array))+
  geom_bar(stat="identity")+
  geom_errorbar((aes(ymin=mean-se, ymax=mean+se)))
 
#genotype plot (1,2,3)
genotype_plot <- ggplot (data=genotype_summarise, aes(x=genotype, y=mean, fill=genotype))+
  geom_bar(stat="identity")+
  geom_errorbar((aes(ymin=mean-se, ymax=mean+se)))

#brick_side plot (Top, N, S, E, W)
side_plot <- ggplot (data=side_summarise, aes(x=side, y=mean, fill=side))+
  geom_bar(stat="identity")+
  geom_errorbar((aes(ymin=mean-se, ymax=mean+se)))

#brick_colour plot (various)
brick_colour_plot <- ggplot (data=brick_colour_summarise, aes(x=brick_colour, y=mean, fill=brick_colour))+
  geom_bar(stat="identity")+
  geom_errorbar((aes(ymin=mean-se, ymax=mean+se)))

#brick_id plot (various)
brick_id_plot <- ggplot (data=brick_id_summarise, aes(x=brick_id, y=mean, fill=brick_id))+
  geom_bar(stat="identity")+
  geom_errorbar((aes(ymin=mean-se, ymax=mean+se)))


# view all plots
array_plot
genotype_plot
side_plot
brick_colour_plot
brick_id_plot

```
  
````{r Further_graphs, echo=TRUE, message=FALSE}

# breakdown further genotype by brick side
gs <-data %>%
  group_by(side, genotype) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))
gs

gs_plot <- ggplot (data=gs, aes(x=side, y=mean, fill=genotype))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
gs_plot

#side by genotype
sg_plot <- ggplot (data=gs, aes(x=genotype, y=mean, fill=side))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
sg_plot

# genotype by array
ga<-data %>%
  group_by(array, genotype) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))
ga

ga_plot <- ggplot (data=ga, aes(x=array, y=mean, fill=genotype))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
ga_plot

#array by genotype
ag_plot <- ggplot (data=ga, aes(x=genotype, y=mean, fill=array))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
ag_plot


# array by side

as<-data %>%
  group_by(array, side) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))
as

as_plot <- ggplot (data=as, aes(x=array, y=mean, fill=side))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
as_plot

# side by array
sa_plot <- ggplot (data=as, aes(x=side, y=mean, fill=array))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
sa_plot

# brick colour and genotype

bg<-data %>%
  group_by(brick_colour,genotype) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))
bg

bg_plot <- ggplot (data=bg, aes(x=brick_colour, y=mean, fill=genotype))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
bg_plot

#genotype and brick colour
gb_plot <- ggplot (data=bg, aes(x=genotype, y=mean, fill=brick_colour))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
gb_plot

# brick colour and array

ba<-data %>%
  group_by(brick_colour,array) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))
ba

ba_plot <- ggplot (data=ba, aes(x=array, y=mean, fill=brick_colour))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
ba_plot

#array and brick colour

ab_plot <- ggplot (data=ba, aes(x=brick_colour, y=mean, fill=array))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
ab_plot

# brick colour and side

bs<-data %>%
  group_by(brick_colour,side) %>%
 summarise(mean=mean(growth, na.rm = TRUE), sd=sd(growth, na.rm = TRUE),n=n(),se=sd/sqrt(n))
bs

bs_plot <- ggplot (data=bs, aes(x=side, y=mean, fill=brick_colour))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
bs_plot

# side and brick colour
sb_plot <- ggplot (data=bs, aes(x=brick_colour, y=mean, fill=side))+
  geom_bar(stat="identity", color="black", position=position_dodge()) +
  geom_errorbar (aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))
sb_plot

#Brick ID-scatter plots
# No side included
brick_id_point<-data %>%
  group_by(brick_id, genotype, array) %>%
  summarise(mean=mean(growth, na.rm = TRUE))
brick_id_point

brick_id_Pplot_geno <- ggplot (data=brick_id_point, aes(x=brick_id, y=mean, colour=genotype, fill=brick_id))+
  geom_point(stat="identity")
brick_id_Pplot_geno

brick_id_Pplot_array <- ggplot (data=brick_id_point, aes(x=brick_id, y=mean, colour=array, fill=brick_id))+
  geom_point(stat="identity")
brick_id_Pplot_array

#side included
brick_id_point_side<-data %>%
  group_by(brick_id, genotype, array, side) %>%
  summarise(mean=mean(growth, na.rm = TRUE))
brick_id_point_side

brick_id_side_plot <- ggplot (data=brick_id_point_side, aes(x=brick_id, y=mean, colour=side, fill=brick_id))+
  geom_point(stat="identity")
brick_id_side_plot

brick_id_geno_Pplot <- ggplot (data=brick_id_point_side, aes(x=brick_id, y=mean, colour=genotype, fill=brick_id))+
  geom_point(stat="identity")
brick_id_geno_Pplot

brick_id_array_Pplot <- ggplot (data=brick_id_point_side, aes(x=brick_id, y=mean, colour=array, fill=brick_id))+
  geom_point(stat="identity")
brick_id_array_Pplot

```
```

```{r Analysis, echo=TRUE, message=FALSE}

#First going to parse out useful columns to new data frame (df)

df<-data %>% 
select(brick_id, genotype, array, side, brick_colour, growth)
head(df)

# scatterplot
pairs(df)

#plot univariate effects of an unbalanced model

require(graphics)
plot.design(growth ~., data=df)

op <- par(mfrow = c(3,1))
with(df, {
interaction.plot(side, array, growth)
interaction.plot(side,genotype, growth)
interaction.plot(genotype, array, growth)
}
)
par(op)

```

  
```

```
  
  
```

```
  
  